---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(xtrial)
```

Let's start with a simple example: a cross-trial causal comparison problem
where we want to estimate the effect of treatment `a` versus treatment `b`
when the two randomized trials do not share the same active comparator.

The `{xtrial}` package provides a common interface for fitting several
estimators under different identification strategies. These include:

* inverse probability weighting (IPW),
* g-computation,
* one-step / efficient influence function (EIF) estimation, and
* targeted minimum loss-based estimation (TMLE).

The package also supports several comparison strategies, including traditional
indirect comparison, observational comparison, direct pooled IPD comparison,
and indirect pooled IPD comparison.

## Simulating Example Data

For illustration, we will use a simulated dataset generated by
`simulate_xtrial_data()`. The data-generating mechanism is motivated by
cross-trial tuberculosis treatment comparisons.

```{r simulate-data}
dat <- simulate_xtrial_data(
  n = 1000
)

head(dat)
```

The simulated data contain the outcome, treatment assignment, study indicator,
and baseline covariates needed by the estimators in `{xtrial}`.

## A First Analysis

The main user-facing function is `xtrial()`. A simple call fits one
comparison approach with one estimator.

For example, the following code fits a TMLE estimator using the direct pooled
IPD comparison approach.

```{r first-fit}
fit <- xtrial(
  data = dat,
  approach = "direct",
  estimator = "tmle",
  nuisance_type = "simple"
)

fit
```

Notice what `xtrial()` returns: an object containing the point estimate,
standard error, confidence interval, and information about the fitted approach
and estimator.

We can store the fitted object and summarize it:

```{r summarize-fit}
summary(fit)
```

## Reusing Arguments

When fitting several estimators to the same dataset, it is often helpful to
store repeated arguments first. This keeps the calls short and makes the
analysis easier to read.

```{r reusable-arguments}
comparison_data <- dat
comparison_approach <- "direct"

fit_tmle <- xtrial(
  data = comparison_data,
  approach = comparison_approach,
  estimator = "tmle",
  nuisance_type = "simple"
)

fit_tmle
```

## Trying Different Estimators

We can compare several estimators under the same identification strategy.

```{r compare-estimators}
fit_ipw <- xtrial(
  data = comparison_data,
  approach = comparison_approach,
  estimator = "ipw",
  nuisance_type = "simple"
)

fit_gcomp <- xtrial(
  data = comparison_data,
  approach = comparison_approach,
  estimator = "gcomp",
  nuisance_type = "simple"
)

fit_onestep <- xtrial(
  data = comparison_data,
  approach = comparison_approach,
  estimator = "onestep",
  nuisance_type = "simple"
)

fit_tmle <- xtrial(
  data = comparison_data,
  approach = comparison_approach,
  estimator = "tmle",
  nuisance_type = "simple"
)
```

The fitted objects can then be printed or summarized individually.

```{r print-estimator-results}
fit_ipw
fit_gcomp
fit_onestep
fit_tmle
```

## Trying Different Comparison Approaches

The same estimator can also be used with different comparison approaches.

```{r compare-approaches}
fit_traditional <- xtrial(
  data = comparison_data,
  approach = "traditional",
  estimator = "tmle",
  nuisance_type = "simple"
)

fit_observational <- xtrial(
  data = comparison_data,
  approach = "observational",
  estimator = "tmle",
  nuisance_type = "simple"
)

fit_direct <- xtrial(
  data = comparison_data,
  approach = "direct",
  estimator = "tmle",
  nuisance_type = "simple"
)

fit_indirect <- xtrial(
  data = comparison_data,
  approach = "indirect",
  estimator = "tmle",
  nuisance_type = "simple"
)
```

```{r print-approach-results}
fit_traditional
fit_observational
fit_direct
fit_indirect
```

## Simple and Flexible Nuisance Estimation

Some estimators require nuisance functions, such as outcome regressions,
treatment mechanisms, or study membership probabilities. The argument
`nuisance_type` controls how these nuisance functions are estimated.

For a fast, reproducible example, we use `nuisance_type = "simple"` above.
When flexible estimation is desired, use:

```{r flexible-nuisance, eval = FALSE}
fit_flexible <- xtrial(
  data = comparison_data,
  approach = "direct",
  estimator = "tmle",
  nuisance_type = "flexible"
)

fit_flexible
```

Flexible nuisance estimation may be slower, but it can be useful when the
working models used by the simple implementation are too restrictive.

## A Minimal Workflow

A typical `{xtrial}` workflow is:

1. prepare or simulate a dataset with the required outcome, treatment, study,
   and covariate variables;
2. choose an identification approach;
3. choose an estimator;
4. call `xtrial()`; and
5. inspect the estimate, standard error, and confidence interval.

For example:

```{r minimal-workflow}
dat <- simulate_xtrial_data(n = 1000)

fit <- xtrial(
  data = dat,
  approach = "direct",
  estimator = "tmle",
  nuisance_type = "simple"
)

summary(fit)
```

## Session Information

```{r session-info}
sessionInfo()
```
